---
title: "Vignette Title"
author: "Vignette Author"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Vignette Title}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

## Abstract
We provide one example for using FLASH, and implementing the greedy and backfitting algorithms. For the speed, we will just just run the backfitting algorithm with a few runs of FLASH during each iteration. We demonstrate that FLASH performs competitively against a positive matrix decomposition (PMD).

## Simulations

### Rank One FLASH

```{r,message=FALSE,warning=FALSE}
library("flashr")
sim_K = function(K, N, P, SF, SL, signal,noise){
  E = matrix(rnorm(N*P,0,noise),nrow=N)
  Y = E
  L_true = array(0, dim = c(N,K))
  F_true = array(0, dim = c(P,K))
  
  for(k in 1:K){
    lstart = rnorm(N, 0, signal*(k/K))
    fstart = rnorm(P, 0, signal*(k/K))
    
    index = sample(seq(1:N),(N*SL))
    lstart[index] = 0
    index = sample(seq(1:P),(P*SF))
    fstart[index] = 0
    
    L_true[,k] = lstart
    F_true[,k] = fstart
    
    Y = Y + lstart %*% t(fstart)
  }
  return(list(Y = Y, L_true = L_true, F_true = F_true, Error = E))
}


sim_hd = function(N, P, SF, SL, signal, a = rchisq(N,3),b = rchisq(P,1),mu = 0){
  
  E = matrix(rep(0,N*P),nrow=N)
  sig2_true = matrix(rep(0,N*P),nrow=N)
  for(i in 1:N){
    for(j in 1:P){
      sig2_true[i,j] = mu + a[i] * b[j]
      E[i,j] = rnorm(1,0,sqrt(mu + a[i] * b[j]))
    }
  }
  
  K=1
  lstart = rnorm(N, 0, signal)
  
  fstart = rnorm(P, 0, signal)
  
  index = sample(seq(1:N),(N*SL))
  lstart[index] = 0
  index = sample(seq(1:P),(P*SF))
  fstart[index] = 0
  
  Y = lstart %*% t(fstart) + E
  
  return(list(Y = Y, L_true = lstart, F_true = fstart, Error = E,sig2_true = sig2_true))
}

```

We have several choices for the variance structure and the factors as well.

Here we provide some example:

#### constant variance 

```{r}
set.seed(99)
N = 100
P = 200
data = sim_K(K=1,N, P, SF = 0.8, SL = 0.5, signal = 1,noise = 0.5)
Y = data$Y
E = data$Error

gf_1 = flash(Y,objtype = "l")
gf_1$sigmae2
# fixed factor
gf_2 = flash(Y,factor_value = data$F_true, fix_factor = TRUE)
gf_2$sigmae2
sum(gf_2$f != data$F_true)
# known variance
gf_3 = flash(Y,objtype = "l", sigmae2_true = 0.25,partype = "known")
gf_3$sigmae2
```

Missing value case

```{r}
set.seed(99)
N = 100
P = 200
data = sim_K(K=1,N, P, SF = 0.8, SL = 0.5, signal = 1,noise = 0.5)
Y = data$Y
Y[1:10,1:10] = NA
E = data$Error

gf = flash(Y,objtype = "l")
gf$sigmae2
```

#### columnwise variance

```{r}
set.seed(99)
N = 100
P = 200
data = sim_hd(N, P, SF = 0.5, SL=0.5, signal =1, a = rep(1,N), b = rchisq(P,1), mu = 0)
Y = data$Y
E = data$Error

gf_1 = flash(Y,objtype = "l", partype = "var_col")
plot(data$sig2_true[1,], gf_1$sigmae2)
# fixed factor
gf_2 = flash(Y, partype = "var_col", factor_value = data$F_true, fix_factor = TRUE)
sum(gf_2$f != data$F_true)
plot(data$sig2_true[1,], gf_2$sigmae2)
# known variance 
gf_3 = flash(Y,objtype = "l", partype = "known",sigmae2_true = data$sig2_true[1,])
plot(data$sig2_true[1,], gf_3$sigmae2)

```

#### Kronecker product variance

```{r}
set.seed(99)
N = 100
P = 200
data = sim_hd(N, P, SF = 0.5, SL=0.5, signal =1, a = rchisq(N,3),b = rchisq(P,1),mu = 0)
Y = data$Y
E = data$Error

gf = flash(Y,objtype = "l", partype = "B")
plot(data$sig2_true,gf$sigmae2)
```

#### Known variance matrix 

```{r}
set.seed(99)
N = 100
P = 200
data = sim_hd(N, P, SF = 0.5, SL=0.5, signal =1, a = rchisq(N,3),b = rchisq(P,1),mu = 0)
Y = data$Y
E = data$Error

gf = flash(Y,objtype = "l", partype = "known",sigmae2_true = data$sig2_true)
plot(data$sig2_true,gf$sigmae2)
```

#### Noisy variance 

```{r}
set.seed(99)
N = 100
P = 200
data = sim_hd(N, P, SF = 0.5, SL=0.5, signal =1, a = rchisq(N,3),b = rchisq(P,1),mu = 0.9)
Y = data$Y
E = data$Error

gf = flash(Y,objtype = "l", partype = "noisy",sigmae2_true = (data$sig2_true - 0.9))
(gf$sigmae2 - (data$sig2_true - 0.9))[1,1]
# missing value 
Y = data$Y
Y[1:10,1:20] = NA
gf = flash(Y,objtype = "l", partype = "noisy",sigmae2_true = (data$sig2_true - 0.9))
(gf$sigmae2 - (data$sig2_true - 0.9))[1,1]
```


### Rank K version 

```{r}
set.seed(99)
N = 100
P = 600
data = sim_K(K=10,N, P, SF = 0.9, SL = 0.8, signal = 1,noise = 1)
Y = data$Y
E = data$Error
plot(svd(Y)$d)
ggd = greedy(Y,K = 100)
#eigen_plot(ggd,sum(Y^2))
gbf <- backfitting(Y, ggd, maxiter_bf = 50)
```
