---
title: "flash"
author: "Wei Wang <<weidong@uchicago.edu>>,  Matthew Stephens <<mstephens@uchicago.edu>> "
date: "`r Sys.Date()`"
# bibliography: flashr.bib
output:
  rmarkdown::html_vignette:
    toc: true
    number_sections: true
vignette: >
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteIndexEntry{flashr}
---

One example for flash and how to use greedy and backfitting. For the speed, I just use backfitting with few runs of flash in each iteration. I also compare with PMD, these are all better than PMD.

```{r}
library("flashr")
sim_K = function(K, N, P, SF, SL, signal,noise){
  E = matrix(rnorm(N*P,0,noise),nrow=N)
  Y = E
  L_true = array(0, dim = c(N,K))
  F_true = array(0, dim = c(P,K))
  
  for(k in 1:K){
    lstart = rnorm(N, 0, signal)
    fstart = rnorm(P, 0, signal)
    
    index = sample(seq(1:N),(N*SL))
    lstart[index] = 0
    index = sample(seq(1:P),(P*SF))
    fstart[index] = 0
    
    L_true[,k] = lstart
    F_true[,k] = fstart
    
    Y = Y + lstart %*% t(fstart)
  }
  return(list(Y = Y, L_true = L_true, F_true = F_true, Error = E))
}

data = sim_K(K=10,N=100, P=300, SF = 0.95, SL = 0.8, signal = 1,noise = 1)
Y = data$Y
L_true = data$L_true
F_true = data$F_true
E = data$Error
svdl = svd(Y)$u
svdf = svd(Y)$v
svdd = svd(Y)$d
plot(svdd)
svdK = 10
sqrt(mean(((Y  - svdl[,1:svdK]%*% diag(svdd[1:svdK]) %*% t(svdf[,1:svdK]) )- E)^2)) / sqrt(mean(((Y - 0)- E)^2))
#sqrt(mean(((Y  - svdd[1] * svdl[,1] %*%  t(svdf[,1]) )- E)^2)) / sqrt(mean(((Y - 0)- E)^2))
ggreedy = greedy(Y,K=30)
# initial gl and gf as svd
gl = ggreedy$l
gf = ggreedy$f
dim(gl)
sqrt(mean(((Y  - gl%*%t(gf) )- E)^2)) / sqrt(mean(((Y - 0)- E)^2))
# proportion in 
eigenvalue = rep(0,dim(gl)[2])
for(i in 1:length(eigenvalue)){
  eigenvalue[i] = (sqrt(mean((gl[,i]%*%t(gf[,i]))^2)))
}
eigenvalue = eigenvalue / sqrt(mean(Y^2))
plot(eigenvalue)


gback = backfitting(Y,gl,gf,tautol = 100, numtau = 5)
gbl = gback$l
gbf = gback$f
dim(gbl)
sqrt(mean(((Y  - gbl%*%t(gbf) )- E)^2)) / sqrt(mean(((Y - 0)- E)^2))
```

The comparing criterion is 
$$\frac{RMSE - RMSE_{opt}}{RMSE_{naive}}$$

If we know the variance, we can do better based on our model.

we can see that if we known the variance matrix, we can do better than using flash with constant variance. I have also compared those with PMD, and the result is flash_hd better then flash which is better then PMD.

```{r}
sim_hd = function(N, P, SF, SL, signal, a = rchisq(N,3),b = rchisq(P,1),mu = 0){

  E = matrix(rep(0,N*P),nrow=N)
  sig2_true = matrix(rep(0,N*P),nrow=N)
  for(i in 1:N){
    for(j in 1:P){
      sig2_true[i,j] = mu + a[i] + b[j]
      E[i,j] = rnorm(1,0,sqrt(mu + a[i] + b[j]))
    }
  }
  
  K=1
  lstart = rnorm(N, 0, signal)
  
  fstart = rnorm(P, 0, signal)
  
  index = sample(seq(1:N),(N*SL))
  lstart[index] = 0
  index = sample(seq(1:P),(P*SF))
  fstart[index] = 0
  
  Y = lstart %*% t(fstart) + E
  
  return(list(Y = Y, L_true = lstart, F_true = fstart, Error = E,sig2_true = sig2_true))
}
library("flashr")
N = 100
P = 200
SF = 0.5
SL = 0.5
signal = 1
data = sim_hd(N, P, SF, SL, signal, a = rchisq(N,3),b = rchisq(P,1),mu = 0)
sigmae2_true = data$sig2_true
Y = data$Y
E = data$Error
ghd = flash_hd(Y,partype = "known",sigmae2 = sigmae2_true)
l = ghd$l
f = ghd$f
sqrt(mean(((Y - l %*% t(f))-E)^2))/sqrt(mean((Y - E)^2))
ghd = flash_hd(Y,partype = "constant")
l = ghd$l
f = ghd$f
sqrt(mean(((Y - l %*% t(f))-E)^2))/sqrt(mean((Y - E)^2))
gvem = flash(Y)
l = gvem$l
f = gvem$f
sqrt(mean(((Y - l %*% t(f))-E)^2))/sqrt(mean((Y - E)^2))
```
